---
title: "assignemnt3_task3_jack_miller"
author: "Jack Miller"
date: "10/23/2021"
output: html_document
---
## Set-up
```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
```

## Read in the Data 
```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names
```


# Section 1 

### Data Visualization 
Histogram of the of the distribution of sizes between purple and red urchins.

```{r}
ggplot(data = urchins,
       aes(x = size)) +
  geom_histogram(bins = 10, aes(fill = common_name)) +
  scale_fill_manual(values = c("purple", "red")) +
theme_minimal() + 
  labs(x = "Size (cm)",
       y = "Number",
       title = "Size Distribution of Red and Purple Urchins",
       fill = "Key") +
  facet_wrap(~ common_name)
```

**Figure 1. This is a histogram of the size distribution of red and purple urchins, taken from 5 sampling sites along the Santa Barbara coast.**

## Descriptive Statistics Table 
Summary statistics table (mean, median, standard deviation, sample size)

```{r}
summary_statistics <- urchins %>% 
  group_by(common_name) %>% 
summarize(mean_size = mean(size),
          max_size = max(size),
          standard_deviation = sd(size), 
          sample_size = n())

final_table <- summary_statistics
  kable(caption = "Table 1. Statistics for Sea Urchin Data"
        digits = c(2,2,2,2,2)
        col.names = c("Species",
                      "Mean body size (cm)",
                      "Max body size (cm)",
                      "Standard deviation of body size",
                      "Sample size (n)"))

```

